---
title: "TimeSeries-BasicHousekeeping"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

### Machine Learning Basic Housekeeping for Time Series Analysis

We will revisit some basic R Syntax for quick recap before we attempt Time Series Forecasting


### How to Load data from various file formats
```{r Reading from various file formats}
# read from CSV
resp1 <- read.csv("/Volumes/Data/CodeMagic/Data Files/Udemy/TimeSeriesData/Resp1.csv",header = TRUE)
# Check some initial rows
head(resp1)
# Learn more about the structure of the data
str(resp1)

# read another file
# This has a separator as ';'
winer1 <- read.csv("/Volumes/Data/CodeMagic/Data Files/Udemy/TimeSeriesData/winequality-red.csv",sep = ";",header = TRUE)
head(winer1)
str(winer1)
# Run a summary statistics
summary(winer1)

# read from xls
library(readxl)
dfb <- read_excel("/Volumes/Data/CodeMagic/Data Files/Udemy/TimeSeriesData/boston1.xls")
head(dfb)
summary(dfb)

# using Rcurl to read csv data hosted online
library(RCurl)
data1 <- read.csv(text=getURL("https://raw.githubusercontent.com/sciruela/Happiness-Salaries/master/data.csv"))
head(data1)
summary(data1)

# Visualize the patterns of NA


```

## Removal & Management of missing values

```{r removal and management of missing values}
library(MASS)

# using inbuilt datasets
data("airquality")
str(airquality)

head(airquality)
str(airquality)

# Number of missing values by columm
colSums(is.na(airquality))

# remove the rows containing NA
aq <- na.omit(airquality)
head(aq)
summary(aq)
str(aq)

# If we want to retain rows which are non-NA
aq2 <- airquality[complete.cases(airquality),]
head(aq2)
summary(aq2)
str(aq2)

# if we replace NAs with 0
aqty <- airquality
aqty[is.na(aqty)] <- 0
head(aqty)
summary(aqty)
str(aqty)

# replace missing values with average values
meanOzone <- mean(airquality$Ozone,na.rm = TRUE)
meanSolar.R <- mean(airquality$Solar.R,na.rm = TRUE)
aqty$Ozone <- ifelse(is.na(airquality$Ozone),meanOzone,airquality$Ozone)
aqty$Solar.R <- ifelse(is.na(airquality$Solar.R),meanSolar.R,airquality$Solar.R)
head(aqty)
print("Before Mean imputation - Summary for Ozone variable")
summary(airquality$Ozone)
print("After Mean imputation - Summary for Ozone variable")
summary(aqty$Ozone)
str(aqty)
```
```{r Visualization of NAs and more statistical imputations}
# Visualize the patterns of NAs
library(mice)
aqty2 <- airquality
md.pattern(aqty2)

# The below visualization shows 72.5% observations in the entire data have no missing values
# 22.9%  missing values in Ozone
library(VIM)
mp <- aggr(aqty2,col = c('navyblue','yellow'),numbers = TRUE, sortVars = TRUE, labels = names(aqty2),
           cex.axis = 0.7, gap = 3, ylab = c("Missing Data","Pattern"))

# Another approach is predictive mean mapping for imputation - we will do 50 iterations and generate 5 datasets
# we will not print the console output in the markdown file
```
```{r,results='hide'}
im_aqty <- mice(aqty2, m=5, maxit = 50, method = 'pmm',seed = 500)
```
```{r}
summary(im_aqty)


# Values imputed in ozone
im_aqty$imp$Ozone

# using the first dataset, we  complete the data
completedData <- complete(im_aqty,1)
head(completedData)
str(completedData)

```

## More Data Cleaning
```{r data cleaning techniques}
# load the dataset
library(RCurl)
eco <- read.csv(text = getURL("https://raw.githubusercontent.com/Jojo666/Sentimental-Analysis_R/master/countries_ecologicalF.csv"))
head(eco)
str(eco)

# GDP.per.Capita column is a factor and has '$' sign.. which is essentially a currency column (numeric)
# We will remove those and create a new column

eco$GDP = as.numeric(gsub("[\\$,]","",eco$GDP.per.Capita))
# Check whether the new column got created
names(eco)
head(eco)

# Check how many columns are there in the dataset
ncol(eco)
# I will subset and work with only few columns
mdf = names(eco) %in% c("Population..millions","GDP","Grazing.Footprint","HDI",
                        "Carbon.Footprint","Cropland","Forest.Land","Urban.Land",
                        "Total.Biocapacity")

# create the subset data
newdata = eco[mdf]
head(newdata)

# If i have to exclude all of the above columns and subset the data
x = eco[!mdf]
head(x)

# Check if we have NA in our newdata
colSums(is.na(newdata))

```

## Including Plots

Perform some Visualizations

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
